---
title: "Air Quality Analysis Report"
author: "Youwei Hu"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

# Introduction

The **Air Quality Dataset** from the UCI Machine Learning Repository contains initially contains 9471 hourly measurements collected in an Italian city from March 2004 to February 2005. It includes concentrations of key air pollutants (e.g., tin oxide, titania, tungsten oxides, benzene) and meteorological conditions (temperature, relative humidity, absolute humidity).

**Analysis Objectives:**

- **Data Cleaning:** Handle missing values and standardize numeric variables.

- **Exploratory Analysis:** Generate descriptive statistics and visualize trends in the data.

- **Model Building:** Split the data into training and testing sets and build linear regression models to predict pollutant levels.

- **Evaluation:** Summarize findings and evaluate model performance.

# Data Cleaning
In this section, I use reproducible code with the here package to load the cleaned dataset (data_clean.rds), which was generated in an R script named clean_data.R. This script performs initial data cleaning, including creating a new datetime variable and removing extra rows and columns that contain only missing values.

```{r, load-data}
here::i_am("final_project.Rmd")
data_clean <- readRDS(
  file = here::here("output/data_clean.rds")
)
str(data_clean)
# Note: The dataset uses ";" as the separator and comma for decimals.
```

# Descriptive Statistics
## Visualization of Correlation 
Correlation matrix plot is performed to visualize the relationships between numeric variables and identify potential multicollinearity issues and guide further analysis.
```{r, correlation_plot}
#|fig.align = "center
#|out.width = "100px"
# Correlation heatmap
knitr::include_graphics(
  here::here("output/corr_plot.png")
)
```
**Correlation Heatmap**

High correlation among pollutant sensor readings suggests that they capture similar environmental trends.
NOx and NO2 levels are strongly related, implying they can be used interchangeably in some models.
Temperature and NOx levels have an inverse relationship, which could be important for air pollution forecasting.
Humidity's role in air pollution appears minimal based on this dataset.

## Tabular analysis

Characteristics of the `r nrow(data_clean)` concentrations from the Air Quality database are displayed in the table below.
```{r, descrptiv_stats}
table_stats <- readRDS(
  file = here::here("output/table_description.rds")
)
table_stats
table_description <- readRDS(
  file = here::here("output/table_stats.rds")
)
table_description
```
**Table Description**

The table presents the summary statistics of key variables in the Air Quality Dataset. It includes various air pollutant concentrations and meteorological factors recorded from a multi-sensor device.

## Time Series Data Visualization
```{r, Data_Visualization}
#|fig.align = "center
#|out.width = "60px"
knitr::include_graphics(
  here::here("output/time_series_plot.png")
)
```
**Time Series Plot for Benzene Concentration**

The time series plot visualizes the fluctuations in benzene concentration (C6H6.GT.) over periods, illustrating trends and seasonal variations.

# Linear Regression Analysis
I split the data into training and testing sets using an 80/20 time-based split and develop a linear regression model to predict benzene concentration (C6H6.GT.).
First, I fit a full model incorporating all regressors. 
```{r, full_model_linear_regression_analysis}
full_model_regression <- readRDS(
  file = here::here("output/primary_regression_table.rds")
) 
full_model_regression
```
Then, I apply stepwise model selection using the Bayesian Information Criterion (BIC) to identify the optimal model.
```{r, bset_model_linear_regression_analysis}
best_model_bic_regression <- readRDS(
  file = here::here("output/best_regression_table.rds")
) 
best_model_bic_regression
```
# Model Evaluation
Finally, I use the selected model to generate predictions on the test data, aiming to achieve the lowest Root Mean Squared Error (RMSE):
```{r model_evaluation}
rmse <- readRDS(
  file = here::here("output/rmse.rds")
)
```
 `r rmse`.

# Conclusion
In this report, we performed an exploratory analysis of the Air Quality Dataset. The steps included:

Data Cleaning: Combining date and time into a datetime variable, removing unused columns, and handling missing values.
Descriptive Analysis: Presenting summary statistics in a formatted table.
Visualization: Creating a time series plot of benzene concentration.
Modeling: Developing a linear regression model to predict benzene levels based on meteorological variables.

